%\documentclass[pt11,a4paper,twoside,reqno,openright]{paper}
%\usepackage[latin1]{inputenc}
%\usepackage{amsmath}
%\usepackage{amsfonts}
%\usepackage{amssymb}
%\usepackage[english]{babel}
%\usepackage{subfigure}
%\usepackage{graphicx}
%\usepackage{float}
%\usepackage{listings,lstautogobble}
%\usepackage[T1]{fontenc}

%\begin{document}

%\noindent {\huge Shapley value and power indices}

%!TeX root = ./GameTheoryNotes.tex

\noindent The core solution is a multifunction from the set of the game to 
real values. Nucleulus is not a multifunction: it attaches 1! value to the 
vectors of all the players.

\noindent The Shapley value is another solution concept; it is a solution 
function, not a multifunction: you only have one distribution of utilities.

\bigskip
\noindent \textbf{Properties of an one-point solution}

\noindent Let $\phi: \mathcal{G}(N) \rightarrow \mathbb{R}^n$ be an one-point solution.

\begin{enumerate}
	\item Efficiency: the sum of the utilities that $\phi$ distributes 
	coincides with $v(N)$, i.e. $\sum_{i \in N}\phi_i(v) = v(N)$.
	
	\item Symmetry: $\phi_i(v) = \phi_j(v) ~\forall i,j$ that are symmetric players, i.e. players s.t. $v(A \cup \{i\}) = v(A \cup \{j\}) ~\forall A \subset N, i,j \notin A$.
	
	\noindent
	\textbf{Example:} in a Parliament, two parties with exactly the same 
	number of seats.

	\item Null-player property: $\phi_i(v) = 0$ for every useless player, i.e. for every player that does not increase (or decrease) 
	the utility of a coalition by entering it (player $i$ is useless if $v(A \cup \{i\}) = v(A) ~\forall A \subset N, i \notin A$).

	\item Additivity: $\phi(v + w) = \phi(v)  + \phi(w)$.
	
	\noindent
	Remember that v and w are vectors and you know how to 
	sum vectors. Problem: we do not actually know what's the meaning of a game 
	which is defined as the sum of two games. We still consider this property 
	because it is useful iot define a solution.
\end{enumerate}

\bigskip
\noindent \textbf{Shapley Theorem:} the function $\sigma: \mathcal{G}(N) \rightarrow \mathbb{R}^n$ defined as
\[
	\sigma_i(v) = \sum_{S \in 2^{N \setminus \{i\}}}{\frac{s!(n-s-1)!}{n!} \left(v(S \cup \{i\}) - v(S)\right)}
\]
is the only functions that satisfies properties (1)-(4).

\bigskip
\noindent If you believe that a function should fulfill all the properties (1)-(4), then you are obliged to consider $\sigma$, since it is the only 
function that satisfies them.

\bigskip
\noindent $m_i(v,S) = v(S \cup \{i\}) - v(S)$ is called the \textbf{marginal contribution} that
player $i$ brings to the coalition $S \cup \{i\}$.

\bigskip
\noindent The Shapley solution for player $i$ is a weighted sum of all the 
marginal contributions of player $i$. 

\bigskip
\noindent What's the meaning of the coefficient? 

\noindent \textbf{Notation:} $s = |S|$: small s is the number of elements in 
the coalition S.

\noindent Suppose that a group of players arrange to meet and suppose that 
the order of arrivals is random. Coefficient: probability that player $i$ 
enters the room and finds there only the players of coalition S. Indeed, the 
number of possible arrival orders is given by all the possible permutations of 
players, i.e. it is $n!$. Player $i$ makes a coalition with S only if all the 
people in S are already there when he enters the room. Of course, the internal 
ordering of players in S does not matter: he will consider all the possible 
permutations of the elements in S. Analogously, also the ordering of all the 
players who will arrive after $i$ does not matter, which means that you 
consider $s!$ and $(n-s-1)!$.\\
$\implies$\\
The coefficient is the probability that player $i$ joins coalition $S$.

\noindent Consider the example of \textbf{slide 11} for the interpretation of 
the coefficient.

\noindent First column: list of all the possible arrival orderings (example: 
first row: they arrive in order "player1 - player2 - player3").

\noindent In the first row, player1 makes a coalition with no one, therefore 
his marginal contribution is $v(\{1\}) = 0$. Player2 makes a coalition with 
player1, therefore his marginal contribution is $v(\{1,2\}) - v(\{1\}) = 4 - 0$. 
Etc.

\noindent Comment: there are repeated values over each column; this is obvious 
because there are multiple situations in which, for example, player1 is the 
first to arrive. 

\noindent All the coalitions that do not contain player1 are:
\[
	\emptyset,~ \{2\},~ \{3\},~ \{2,3\}
\]
therefore marginal values are:
$v(1) - v(\emptyset)$\\
$v(\{1,2\}) - v(2)$\\
$v(\{1,3\}) - v(3)$\\
$v(N) - v(\{2,3\})$.

\bigskip
\noindent \textbf{Proof of Shapley theorem:} the proof is divided in two steps:
\begin{itemize}
	\item We show that the function $\sigma$ satisfies properties (1)-(4):
	\begin{enumerate}
		\item Efficiency: I write in $\sum_{i=1}^n{\sigma_i(v)}$ the definition 
		of $\sigma_i$ given by the Shapley theorem. 
		
		\noindent $v(N)$ appears in the sum 
		only when $S = N \setminus \{i\}$, because the internal sum is defined over $S$ s.t. 
		$S \subseteq (N \setminus \{i\})$, therefore it appears once for every player, with coefficient $\frac{s!(n-s-1)!}{n!} = \frac{1}{n}$. 
		
		\noindent
		We consider any coalition $A \neq N$ and we prove that $v(A)$ appears with coefficient 
		zero (because it appears sometimes with positive and sometimes with negative 
		coefficient). 
		
		\noindent We have positive coefficient when $i$ belongs 
		to $A$ and we set, in the formula, $S = A \setminus \{i\}$. In this case, being $a$ the cardinality of $A$, we have $s = a-1$ $\implies$ the positive coefficient $\frac{(a-1)!(n-a)!}{n!}$ appears $a$ times (once for every player $i \in A$) and therefore we get $\frac{a!(n-a)!}{n!}$.
		
		\noindent
		We have the negative coefficient when player $i$ does not belong to $A$, i.e. 
		it appears $n-a$ times, setting in the formula $S = A$ $\implies$ its total contribution is $- \frac{a!(n-a-1)!}{n!}(n-a) = 
		- \frac{a!(n-a)!}{n!}$.
	
		\item Symmetry: I should prove that $\sigma_i(v) = \sigma_j(v)$. 
		Remember: these are two different sums: in $\sigma_i$ I consider coalitions 
		that do not contain $i$, but of course they may contain $j$ (and viceversa). 
		I proceed by splitting the sum between the coalitions that do not contain 
		neither $i$ nor $j$ and then I add what I need. Therefore, in the second 
		term of $\sigma_i$, for example, the index is S (which does not contain 
		neither $i$ nor $j$), but the actual coalition is $S \cup \{j\}$. 
		With this trick, the two formulas are exactly the same, I only have to 
		interchange $i$ and $j$. Since the union is symmetric and since in each 
		sum I have the same coefficients, I have exactly that $\sigma_i = \sigma_j$, 
		which corresponds to symmetry.
		
		\noindent In mathematical terms,
		\[
			\sigma_i(v) = \sum_{S \in 2^{N \setminus (\{i\} \cup \{j\})}}
			{\frac{s!(n-s-1)!}{n!}\left[v(S \cup \{i\}) - v(S)\right]} +\]
			\[ 
			\sum_{S \in 2^{N \setminus (\{i\} \cup \{j\})}}
			{\frac{(s+1)!(n-s-2)!}{n!}\left[v(S \cup \{i\} \cup \{j\}) - v(S \cup \{j\})\right]}
		\]
		while
		\[
		\sigma_j(v) = \sum_{S \in 2^{N \setminus (\{i\} \cup \{j\})}}
		{\frac{s!(n-s-1)!}{n!}\left[v(S \cup \{j\}) - v(S)\right]} +\]
		\[ 
		\sum_{S \in 2^{N \setminus (\{i\} \cup \{j\})}}
		{\frac{(s+1)!(n-s-2)!}{n!}\left[v(S \cup \{i\} \cup \{j\}) - v(S \cup \{i\})\right]}
		\]
		which are equal for symmetric players.
		
		\item Null player property means that $v(S \cup \{i\}) - v(S)$ is always 
		zero, which of course entails that $\sigma_i$ is always zero (I can have 
		whatever coefficient).
	
		\item Additivity: also here the coefficients do not matter. I only have to 
		apply distributive law (it. raccoglimento a fattor comune).
	\end{enumerate}

	\item We prove that $\sigma$ is unique.
	\noindent Consider the 
	unanimity game, which is a basis for the set of games $\mathcal{G}(N)$. 
	What happens is the following:
	\begin{itemize}
		\item players inside $A$ are symmetric (they are all necessary to make 
		the coalition winning)
	
		\item players outside $A$ are null players, because if $A$ is a winning 
		coalition it has value 1 and this value remains 1 no matter how many 
		players I add.
	\end{itemize}
	Therefore, players outside$A$ must take 0 and players inside A must take all the same value. Moreover, the function is efficient, therefore the value of the game, which is 1, is equally distributed among players in $A$.\\
	$\implies$\\
	The function is uniquely determined over the basis of $\mathcal{G}(N)$ given by unanimity games.
	
	\noindent The same argument can be applied to the game $Cu_A$, where $C$ is a constant in $\mathbb{R}$
	(remember: a game is a vector and I can always multiply a vector by a 
	constant) $\implies$ players outside A must take 
	$\frac{C}{n}$.
	
	\noindent The additivity axiom entails uniqueness.
\end{itemize}

\bigskip
\noindent \textbf{Shapley value for simple games}

\noindent Simple games are the one in which the value of a coalition is either 
0 or 1. In this case, it is easy to compute marginal contributions: they can 
only be given by $(1-1)$, $(0 - 0)$ or $(1 - 0)$. The only important case is the 
last one: to add a player to a winning coalition does not change anything and 
to add a player to a losing coalition does not tell anything interesting if 
the new coalition is still losing. 
The case $(0 - 1)$ cannot occur: if I add a player to a winning coalition, it 
cannot become a losing coalition.

\noindent Iot compute $\sigma_i$, we must count the number of times in which 
player $i$ is crucial for a coalition.

\bigskip
\noindent If we define $\mathcal{A}_i$ as the family of coalitions $A$ s.t. $i \notin A$, $A$ is losing and $A \cup \{i\}$ is winning, then
\[
	\sigma_i(v) = \sum_{A \in \mathcal{A}_i}\frac{a!(n-a-1)!}{n!}
\]

\noindent As an alternative form, if we define $\mathcal{W}_i$ as the family of coalitions $W$ s.t. $i \in W$ and $W \setminus \{i\}$ was winning (of course $W$ is still winning), then
\[
	\sigma_i(v) = \sum_{W \in \mathcal{W}_i}\frac{(w-1)!(n-w)!}{n!}
\]

\bigskip
\noindent \textbf{Example:} airport game, from slide 12 of slide deck "Lesson7".

\noindent The game: player1 only need 1km truck, player2 needs 2km and player3 
needs 3km. The total cost is $c_3$. How to share this cost among the players? 

\noindent For player2, we have:
$v(2) - v(\emptyset) = c_2$\\
$v(\{1,2\}) - v(1) = c_2 - c_1$\\
$v(\{2,3\}) - v(3) = 0$\\
$v(N) - v(\{1,3\}) = 0$\\
The last two differences are zero because if we have player3 we must build 
3km and therefore the arrival of player2 does not change anything (we do not 
have kms to add only because player2 has joint the coalition). 
Therefore $\sigma_2 = \frac{c_2}{3} - \frac{c_2-c_1}{6}$

\bigskip
\noindent \textbf{Slide 13}

\noindent In simple games, the main goal is to divide coalitions in winning 
and losing ones. The interpretation of a solution is not utility: what is 
important is the fact that, for example, if two parties are strong enough to 
get the majority, they have a lot of power. The game is not expressed in terms 
of utility. Shapley values express the fraction of power owned by every player, 
rather than their utility.

\noindent \textbf{Example:} consider two different situations. You have 
four players and each one of them owns a portion of the stocks of a 
company. In the first case, we have that their percentages are:\\
$\{10, 20, 30, 40\}$\\
while in the second case we have: \\
$\{10, 21, 30, 39\}$.

\noindent The two situations appears to be almost the same, but in fact they are 
not: in the first case, sometimes player1 is crucial iot build a winning 
coalition, since a coalition is winning iff it owns more than $50\%$. 
In the second case, actually, player1 is the null player: all the other 
can make winning coalitions without him.

\noindent On the other hand, a third situation\\
$\{11, 20, 30, 39\}$\\
turns out to be almost equal to the first one. The point are not numbers, but 
the real meaning of those numbers.

\bigskip
\noindent \textbf{Remark:} when I'm measuring the power of the players, what 
actually means is the ratio: it is the fact, for example, that my fraction 
of power is twice the yours. In this context, efficiency is not a 
requirement I must have.

\bigskip
\noindent \textbf{Probabilistic power index:} a probabilistic power index $\psi$ on the set of simple games is
\[
	\psi_i(v) = \sum_{S \in 2^{N \setminus \{i\}}}p_i(S)m_i(v,S)
\]
where $p_i$ is a probability measure over $2^{N \setminus \{i\}}$.

\bigskip
\noindent \textbf{Semivalues:} a probabilistic power index $\psi$ on the set of simple games is a semivalue if there exists a vector $(p_0,...,p_{n-1})$ s.t.
\[
	\psi_i(v) = \sum_{S \in 2^{N \setminus \{i\}}}p_sm_i(v,S)
\]

\bigskip
\noindent \textbf{Remark:} since the index is probabilistic, two conditions must hold:
\begin{enumerate}
	\item $p_s \geq 0 ~\forall s$
	\item $\sum_{s=0}^{n-1}\binom{n-1}{s}p_s = 1$
\end{enumerate}
In particular, if $p_s > 0 ~\forall s$, the semivalue is called \textbf{regular}.

\bigskip
\noindent Remember: $s$ is the cardinality of coalition $S$. What I am saying 
is that the probability to make a coalition with coalition A is equal to 
the probability to make a coalition with B, if both A and B have the same 
cardinality. This assumption is not reasonable, for example, in a parliament, 
when I do not make coalitions with some parties only because they have the 
right number of seats {\huge ma certo XD}.

\bigskip
\noindent \textbf{Examples of semivalues:}
\begin{itemize}
	
	\item Shapley value
	
	\item Banzhaf value: it says that the number of players in a coalition 
	does not matter when I have to decide which coalition I want to belong to. 
	Every coalition is equally likely. This is sometimes reasonable, when I 
	know that players aggregate no matter the number they are.
	\[
		\beta_i(v) = \sum_{S \in 2^{N \setminus \{i\}}}
		{\frac{1}{2^{n-1}}m_i(v,S)}
	\]

	\item Dictatorial value: it is used when I have the feeling that people might 
	aggregate in coalitions, but this is unlikely.
	\[
		p_0 = 1, ~p_s = 0 ~\forall s=1:n-1
	\]
\end{itemize}

\bigskip
\noindent Properties of semivalues: if you don't consider Shapley, they fulfill 
all the properties of Shapley but for efficiency.

\bigskip
\noindent \textbf{Example from slide 16:}

\noindent All non-veto players are symmetric and all veto players are 
symmetric $\implies$ I only have to compute one Shapley value and then 
apply symmetry and additivity. Calculations should be equal for non-veto 
players. A non-veto player is crucial if he joins a coalition where we already 
have 5 veto players and 3 non-veto players. To be in this situation, I have 
nine players (the remaining non-veto players when I am considering player 
$i$) and I have to take 3.

\noindent On the other hand, a veto player is crucial when I have 4 veto 
players and at least 4 non-veto players.

\noindent The ratio between the Banzhaf of non-veto and the Banzhaf of 
veto players is almost 10, while for Shapley it is almost 100 $\implies$ 
there is a factor of 10 among them.

%\end{document}
